---
title: "R Notebook"
output: html_notebook
---

Initialization of packages and sourcing the necessary functions (to install on R-server, specify directory for the packages):
```{r}
# exploring the RTCGA package
# https://www.r-bloggers.com/rtcga-factory-of-r-packages-quick-guide/
# a <- tools:::.read_repositories(file.path(R.home("etc"), "repositories"))
# install.packages("BiocInstaller", repos = a["BioCsoft", "URL"], lib = "../r-packages/")
# source("http://bioconductor.org/biocLite.R")
# install.packages("BiocInstaller", repos="https://bioconductor.org/packages/3.2/bioc", lib = "../r-packages/")
# url <- "http://bioconductor.org/packages/3.1/bioc/src/contrib/PACKAGES"
# t <- tempfile()
# download.file(url,t)

# remove.packages("BiocInstaller", lib = "../r-packages/")
# source("../Downloads/biocLite.R")

# biocLite(lib.loc = "../r-packages/")
# biocLite("RTCGA", lib.loc = "../r-packages/")
# biocLite("RTCGA.clinical", lib.loc = "../r-packages/")
# biocLite("RTCGA.mRNA", lib.loc = "../r-packages/")
# biocLite("RTCGA.miRNASeq")
# biocLite("RTCGA.methylation")
# biocLite("RTCGA.CNV")
library("RTCGA", lib.loc = "../r-packages/")
# library("RTCGA.clinical", lib.loc = "../r-packages/")
# library("RTCGA.mRNA", lib.loc = "../r-packages/"))
# library("RTCGA.miRNASeq")
# library("RTCGA.methylation")
# library("RTCGA.CNV")

if (!require(devtools)) {
    install.packages("devtools")
    require(devtools)
}
install_github("RTCGA/RTCGA", build_vignettes = TRUE)
install_github("RTCGA/RTCGA.clinical.20160128", build_vignettes = TRUE)
install_github("RTCGA/RTCGA.mRNA.20160128", build_vignettes = TRUE)
install_github("RTCGA/RTCGA.miRNASeq.20160128", build_vignettes = TRUE)
library("RTCGA.clinical.20160128")
library("RTCGA.mRNA.20160128")
library("RTCGA.miRNASeq.20160128")

# install.packages(lib = "../r-packages/")

source("RTCGA_functions.R")
```
Choose which diseases and omics data sources we want to consider:
```{r}
diseases <- c("BRCA", "GBM", "OV")
omics <- c("mRNA", "miRNASeq")#, "methylation", "CNV")
```
Add the variable "days_to_death" to the omics data (and remove patients that don't have this variable) and find the intersection of patients that are in all data sets:
```{r}
AddSurvDataToOmics(diseases, omics, save = FALSE)

for(dis in diseases){
  assign(sprintf("%s_inter", dis), FindIntersection(omics, dis, save = FALSE))
}
```
There is a problem: for GBM, there is only miRNA-seq data for 5 patients and these do not show up in both the CNV and the methylation data. Therefore there are no patients for which all data is there and the intersection is empty.
```{r}
#install.packages("VennDiagram")
# library("VennDiagram")
GBM_venn <- venn.diagram(list(CNV = GBM_CNV$patient_id, Methylation = GBM_methylation$patient_id, miRNA = GBM_miRNASeq$patient_id), filename = NULL)
grid.newpage(recording = FALSE)
grid.draw(GBM_venn, recording = FALSE)
```

There are 283 GBM patients for which there is methylation and CNV data; I should probably look for mRNA and more miRNA data. First, check if the downloading of the data using the RTCGA package is being done correctly.
```{r}
?RTCGA
# Snapshots of the clinical, mutations, CNVs, rnaseq, RPPA, mRNA, miRNASeq and methylation datasets from the 2015-11-01 release date (check all dates of release with checkTCGA('Dates')) are included in the RTCGA.data family (factory) that contains 9 packages
# I should do that
# checkTCGA('Dates')
setRepositories()
# type: 1 2 3 4 5 6 10
biocLite("RTCGA.mRNA.20160128", lib.loc = "../r-packages/")
# was successful
library("RTCGA.mRNA.20160128", lib.loc = "../r-packages/")
# test <- RTCGA.mRNA.20160128 doesn't work
downloadTCGA("GBM", dataSet = "mRNA", date = "2016-01-28", destDir = ".")
test_mRNA_GBM <- readTCGA("gdac.broadinstitute.org_GBM.mRNA_Preprocess_Median.Level_3.2016012800.0.0/GBM.medianexp.txt", dataType = "mRNA")
```
Using the ready-made RTCGA.xxx packages is not the best idea. So I will have to rewrite the functions to download and edit the data.
```{r}
downloadTCGA(diseases, "Merge_Clinical.Level_1", destDir = "../Data")
?checkTCGA
datasets <- checkTCGA("DataSets", "BRCA")
```
For the newest version of Rstudio, a new RTCGA-package is available with the newest data (2016-01-28). I can install this package on my laptop, but the rstudio on the Linux computer first needs to be updated.
