---
title: "Thinking about how to connect miRNA's to pathways"
output: html_notebook
---

30-03-2017

miRNa's target mRNA's. The predicted miRNA-gene interactions are biased. Read "Pathway analysis from lists of microRNAs: common pitfalls and alternative strategy" and "Integrated gene set analysis for microRNA studies"

```{r}
# biocLite("targetscan.Hs.eg.db")
library("targetscan.Hs.eg.db")
```

03-04-2017

Godard et al show that first looking up miRNA target genes and then doing pathway analysis is biased and is not the best way. They suggest to convert the pathway gene lists to pathway miRNA lists, and do pathway analysis directly on the miRNA data. This prevents miRNA's being 'counted double' since every miRNA will only show up once in a pathway list. But, maybe this double counting is actually biologically relevant. 

Maybe it would be an idea to incorporate something like the correlation between a miRNA and its target gene. Or is there a linear relation between amount of miRNA present and amount of mRNA target present?

For now, the idea is:

- I've got a list of pathways with their respective genes
- I've got a list of miRNA's (the full outer join of the rownames of all miRNA datasets)
- I need a lookup table to connect miRNA's and their target genes / mRNA's (the suggested source mirDIP seems a good choice, it is a combination of several prediction databases)

Downloaded all data from mirDIP:

```{r}
colns <- scan("/home/bit/berg0/mirDIP/mirDIP-All-Data-Version1.0/README.txt", what = character(), sep = "\n")[-1]
gene_to_mir <- read.table("/home/bit/berg0/mirDIP/mirDIP-All-Data-Version1.0/mirDIP-All-Data-Version1.0.txt", sep = "\t", quote = "")
colnames(gene_to_mir) <- colns
```

Now I've also got a lookup table. =D

Next thing to fix is getting a list of pathways with their respective miRNA's that affect it. Approach: take pathway-gene list, per pathway look up targetting miRNA's and put them in a new pathway-miRNA list.

```{r}
GeneTomiRNA <- function(pwgenes, high_conf, lookup = gene_to_mir){
  res <- lookup[lookup$'Gene Symbol' %in%
                  entrez_hgnc[entrez_hgnc$entrezgene %in% pwgenes, 'hgnc_symbol'] & high_conf, 'MicroRNA']
  return(res)
}
vGeneTomiRNA <- Vectorize(GeneTomiRNA)
```


```{r}
high_conf <- gene_to_mir$`Normalized Score` >= 90
pw_miRNA <- lapply(kegg2, GeneTomiRNA, high_conf = high_conf)
```

This doesn't give the desired result...

```{r}
pw2_to_mir <- GeneTomiRNA(kegg2[[2]], high_conf = high_conf)
```

First off all, we're getting a very high number of miRNA's. We should adjust the confidence for a certain interaction.

Done. Now I've got a list of pathways with their respective miRNA's.