---
title: "Getting normally distributed data"
output: html_notebook
---

First checking to see which datasets likeley come from a normal distribution. Sequencing data is definitely not normally distributed, since they are counts.

```{r}
data <- BRCA_mRNA[1,]
normal_dis <- rnorm(ncol(BRCA_mRNA), mean = mean(data), sd = sd(data))
plot(density(data), main = "comparing data to normal distribution"); lines(density(normal_dis), col = "red"); legend("topright", legend = c("data", "normal distribution"), pch = 16, col = c("black", "red"))
```

```{r}
qqnorm(data);qqline(data)
```
```{r}
shapiro.test(data)
```
Based on the Shapiro-Wilk normality test, we can reject the null hypothesis that this example sample comes from a normal distribution. I could do this test for every dataset, however, I would probably find that all the data is not normally distributed.

I'm using the *voom()* function from the *limma* package to convert sequencing data to a normal distribution. For this function, missing values are not allowed so I first need to impute missing values. But first, I need to take a closer look at the miRNAseq data. It turns out, the mRNAseq and miRNAseq data is already log2-transformed to bring it closer to a normal distribution (https://confluence.broadinstitute.org/display/GDAC/Documentation#Documentation-miRseqPipelines). But for miRNAseq, also the old data is saved so I need to extract the transformed data:

```{r}
for(d in names(diseases)){
  obj <- get(sprintf("%s_miRNAseq", d))
  assign(sprintf("%s_miRNAseq2", d), obj[,seq(from = 2, to = ncol(obj), by = 3)])
}
```
```{r}
for(d in names(diseases)){
  assign(sprintf("%s_miRNAseq", d), get(sprintf("%s_miRNAseq2", d)))
}
```

CNV data is also in some sort of log2 transformed form, with the addition that 2 is subtracted from all values so that a value of 0 means no change from the normal copy number of 2. This means all data is already transformed to get it as close as possible to a normal distribution.

Some rows are completely filled with NA's. To correct this:
```{r}
for(d in names(diseases)){
  for(o in omics){
    objname <- sprintf("%s_%s", d, o)
    print(objname)
    if(exists(objname)){
      obj <- get(objname)
      assign(objname, obj[complete.cases(obj),])
    }
  }
}
```


Now we can impute missing values:
```{r}
# biocLite("impute")
library(impute)
```
```{r}
for(d in names(diseases)){
  for(o in omics[-5]){
    objname <- sprintf("%s_%s", d, o)
    if(exists(objname)){
      obj <- get(objname)
      assign(objname, obj[complete.cases(obj),])
      print(d)
      print(o)
      assign(
        sprintf("%s_%s_imputed", d, o),
        impute.knn(obj),
        envir = .GlobalEnv
      )
    }
  }
}
```

```{r}
for(d in names(diseases)){
  for(o in omics[-5]){
    if(exists(sprintf("%s_%s_imputed", d, o))){
      assign(sprintf("%s_%s", d, o), get(sprintf("%s_%s_imputed", d, o))$data)
    }
  }
}
```
```{r}
for(d in names(diseases)){
  objname <- sprintf("%s_pheno", d)
  obj <- get(objname)
  rownames(obj) <- toupper(rownames(obj))
  assign(objname, obj)
}

vpaste0 <- Vectorize(paste0)
rownames(BRCA_pheno) <- unlist(lapply(strsplit(rownames(BRCA_pheno), split = "[-]"), vpaste0, collapse = "."))

for(d in names(diseases)[-1]){
  objname <- sprintf("%s_pheno", d)
  obj <- get(objname)
  rownames(obj) <- unlist(lapply(strsplit(rownames(obj), split = "[-]"), vpaste0, collapse = "."))
  assign(objname, obj)
}
```

```{r}
# library("sva")
for(d in names(diseases)){
  print(d)
  assign(sprintf("%s_mRNA_corrected", d), CorrectBatch(
    get(sprintf("%s_mRNA", d)),
    get(sprintf("%s_mRNA_patients", d)),
    get(sprintf("%s_pheno", d))
  ))
}
```
```{r}
for(d in names(diseases)){
  if(exists(sprintf("%s_mRNA_corrected", d))){
    assign(sprintf("%s_mRNA", d), get(sprintf("%s_mRNA_corrected", d)))
  }
}
```

And now save everything:

```{r}
for(d in names(diseases)){
  print(d)
  if(exists(sprintf("%s_pheno", d))){
    save(list = sprintf("%s_pheno", d), 
         file = sprintf("/home/abidata/Joanna/%s/%s_pheno.rda", d, d))
  }
  for(o in omics){
    print(o)
    if(exists(sprintf("%s_%s", d, o))){
      save(list = sprintf("%s_%s", d, o), 
           file = sprintf("/home/abidata/Joanna/%s/%s/%s_%s.rda", d, o, d, o))
      save(list = sprintf("%s_%s_patients", d, o),
           file = sprintf("/home/abidata/Joanna/%s/%s/%s_%s_patients.rda", d, o, d, o))
    }
  }
}
```

